---
title: "P8105_HW6_ah4028"
author: "Austin Hsin"
date: "2023-11-28"
output: html_document
---

```{r setup, include=FALSE}
#Setup code including libraries and plot colors and theme settings
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 2

First, we will import the NOAA weather data provided by the homework document.

```{r weather, message = FALSE}
#Importing NOAA Weather dataframe
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Next, we will bootstrap our linear model of tmax = tmin + prcp and estimate r^2 and log(β^1 ∗ β^2) from the samples.

```{r bootstrap, warning = FALSE}
#Bootstrapping weather_model dataframe
weather_model_df <- 
  modelr::bootstrap(weather_df, n = 5000) |> #5000 samples
  mutate(
    models = map(strap, \(weather_df) lm(tmax ~ tmin + prcp, data = weather_df)), #linear model of tmax
    results = map(models, broom::tidy),
    model_glance = map(models, broom::glance)) |> #fitted regression statistics, specifically for r^2
  select(-strap, -models) |> #removing unnecessary variables
  unnest(results, model_glance) |> 
  select(.id, term, estimate, r.squared) |> #selecting relevant variables
  pivot_wider(names_from = term,
              values_from = estimate) |>
  mutate(beta_calc = tmin + prcp) #log(β^1∗β^2) calculation 
```

Next, we will calculate the 95% confidence intervals for r^2 and log(β^1 ∗ β^2).

```{r calculations}
#95% Confidence Interval for r^2
weather_model_df |>
  summarize(
    r_sq_ci_lower = quantile(r.squared, 0.025), 
    r_sq_ci_upper = quantile(r.squared, 0.975))

#log(β^1∗β^2) and 95% confidence interval
weather_model_df |>
  summarize(
    beta_ci_lower = quantile(beta_calc, 0.025),
    beta_ci_upper = quantile(beta_calc, 0.975))
```

The distribution of r^2 has a 95% confidence interval of (0.8882, 0.9409) and
the distribution of log(β^1 ∗ β^2) has a 95% confidence interval of (0.9798, 1.044).

Finally, we will plot the distribution of the estimates.

```{r bootstrap plots}
#r^2 distribution
weather_model_df |>
  ggplot(aes(x = r.squared)) + geom_density() +
  labs(title = "Distribution of r^2 estimates from bootstrap samples of NOAA tmax weather model") + #title
  scale_x_continuous(n.breaks = 10) #ensuring adequate x axis breaks

#log(β^1∗β^2) distribution
weather_model_df |>
  ggplot(aes(x = beta_calc)) + geom_density() +
  labs(title = "Distribution of beta_calc from bootstrap samples of NOAA tmax weather model") + #title
  scale_x_continuous(n.breaks = 11) #ensuring adequate x axis breaks
```

From the plot of the distribution of `r.squared`, we can see that the distribution is left skewed between 0.861 to 0.952 with a central tendency at 0.916.
From the plot of the distribution of `beta_calc`, representing log(β^1 ∗ β^2), we can see that the distribution is symmetrical between 0.952 and 1.073 with a central tendency at 1.014.

## Problem 3

First, we will import and tidy the `birthweight.csv` data.

```{r birthweight}
birthweight_df = read_csv("./data/birthweight.csv", 
                          col_names = TRUE, #keeping column names
                          col_types = "fnnnnnnnnfnnnnnnnnnnn") |> #setting column types
  janitor::clean_names()
```

Description of proposed model thought process
I included `babysex` and `mrace` as controls, just instinctively given epi background.
Baby measurements `bhead` and `blength` are self-evidently, likely to have positive association.
Mother attributes wise, I included `gaweeks`, `mheight`, `momage`, `delwt`, and `wtgain` as physical attributes as well as `smoken` and `fincome` as external attributes.


